Exploring Bounds

### 2015-04-23 draft.tex
started new lexicode proof

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfig}
\usepackage{enumitem}

@@

Notice that this is simply the even weight vectors.  In fact, it is true that for any $q$

THIS IS THE EVEN WEIGHT CODE! prove it.

\begin{enumerate}[start=0]
\item $\mathbf{c}_0$ is the trivial linear combination.
\item $\mathbf{c}_1 = \mathbf{c}_1$
\item $\mathbf{c}_2 = \mathbf{c}_2$
\item $\mathbf{c}_3 = \mathbf{c}_1 + \mathbf{c}_2$.  $00110 = 00011+ 00101$.
\item $\mathbf{c}_4 = 1\cdot\mathbf{c}_4$
\item $\mathbf{c}_5 = \mathbf{c}_1 + \mathbf{c}_4$. $01010 = 00011 + 01001$.
\item $\mathbf{c}_6 = \mathbf{c}_2 + \mathbf{c}_4$. $011000 = 00101 + 01001$.
\item $\mathbf{c}_7 = \mathbf{c}_3 + \mathbf{c}_4 = \mathbf{c}_1 + \mathbf{c}_2 + \mathbf{c}_4$. $01111 = 00011 + 00101+01001$.
\end{enumerate}

@@

Consequently, it is enough to show that any code $\mathcal{C}$ with covering radius $d-1$ or less meets the Gilbert bound.  Recall that the Gilbert bound can be thought of as a restatement of the 
following fact: (number of spheres) $\cdot$ (number of vectors in a sphere with radius $d-1$) $\ge$ (number of vectors in $\mathbb{F}_q^n$).  Since our covering radius
is $d-1$ or less, and we have that all codewords are distance $d$ or more apart, consequently no sphere of radius $d-1$ in $\mathbb{F}_q^n$ can contain more than one codeword.
As a result, 
Hence, $\mathcal{C}$ meets the Gilbert bound.
\end{proof}

Some notable codes we have discussed in class are in fact lexicodes. For example, the binary lexicodes with $d$=3 and $n=2^m-1$ are the binary hamming codes.  It can be quite
tedious to compute and confirm something like this by hand, so I wrote a short C++ program that implements the Lexicode algorithm as described above.

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{verbatim}
\usepackage{graphicx}
%\usepackage{caption}
%\usepackage{subfig}
%\usepackage{enumitem}

@@

Notice that this is simply the even weight vectors.  In fact, it is true that
for any $q$.
THIS IS THE EVEN WEIGHT CODE! prove it.

\begin{enumerate}%[start=0]
\item $\mathbf{c}_0$ is the trivial linear combination.
\item $\mathbf{c}_1 = \mathbf{c}_1$
\item $\mathbf{c}_2 = \mathbf{c}_2$
\item $\mathbf{c}_3 = \mathbf{c}_1 + \mathbf{c}_2$.  $00110 = 00011+ 00101$.
\item $\mathbf{c}_4 = 1\cdot\mathbf{c}_4$
\item $\mathbf{c}_5 = \mathbf{c}_1 + \mathbf{c}_4$. $01010 = 00011 + 01001$.
\item $\mathbf{c}_6 = \mathbf{c}_2 + \mathbf{c}_4$. $011000 = 00101 + 01001$.
\item $\mathbf{c}_7 = \mathbf{c}_3 + \mathbf{c}_4 = \mathbf{c}_1 + \mathbf{c}_2 + \mathbf{c}_4$. $01111 = 00011 + 00101+01001$.
\end{enumerate}

@@

Consequently, it is enough to show that any code $\mathcal{C}$ with covering radius $d-1$ or less meets the Gilbert bound.  Recall that the Gilbert bound can be thought of as a restatement of the 
following fact: (number of spheres) $\cdot$ (number of vectors in a sphere with radius $d-1$) $\ge$ (number of vectors in $\mathbb{F}_q^n$).  Since our covering radius
is $d-1$ or less, and we have that all codewords are distance $d$ or more apart, consequently no sphere of radius $d-1$ in $\mathbb{F}_q^n$ can contain more than one codeword.
Our spheres will still overlap, however, meaning (number of vectors in
$\mathbb{F}_q^n$)/(number of vectors in a sphere with radius $d-1$) will
certainly be greater than or equal to the number of spheres around codewords.
Hence, $\mathcal{C}$ meets the Gilbert bound.
\end{proof}

Some notable codes we have discussed in class are in fact lexicodes.
[REFERENCE]

We can prove a basic property of binary Lexicodes.  First, though, we
need an important definition.

\begin{defn}
The $\mathbf{nim-sum}$ NS($v$) of a vector $v \in \mathbb{F}_2^n$ is the sum of
it's coordinates in GF($2$).
\end{defn}

\begin{defn}
A vector $v\in \mathbb{F}_2^n$ is $\mathbf{zero-sum}$ if NS($v$)=0.
\end{defn}

We have used the idea of nim-sums and zero-sums throughout the course, but instead we have
referred to this notion in terms being even-like or odd-like.  This is the
exact same notion, only in different terms.

\begin{claim}
For $d=q=2$, $\forall v\in \mathcal{L}$, NS($v$) $= 0$, and hence $\mathcal{L}$ is an even-like code.
\end{claim}

\begin{proof}
We will prove this by inducting on $n$.  In the base case of $n=1$,
$\mathcal{L}=\{0\}$, which is trivially zero-sum.  Assume that for the
lexicode of length $n$, which we will denote $\mathcal{L}(n)$, all codewords are zero-sum.  Now, we must show that all
the codewords in $\mathcal{L}(n+1)$ are zero-sum as well.  First, note that all
codewords in $\mathcal{L}(n)$ are also in $\mathcal{L}(n+1)$, except with an
extra leading zero.  By our assumption, these codewords are still zero-sum, as
adding a zero in front will not change the nim-sum calculation.
Now, we must show that any vector we can add to this set as proscribed by our
algorithm must necessarily be a zero-sum vector as well.
\end{proof}

For example, the binary lexicodes with $d$=3 and $n=2^m-1$ are the binary hamming codes.  It can be quite
tedious to compute and confirm something like this by hand, so I wrote a short C++ program that implements the Lexicode algorithm as described above.
